Finding a closest point in a lattice of Voronoi's first kind

### 2014-01-13 paper.tex
added stuff

\begin{proposition}\label{eq:latticepointsinsphere}
Let $\Lambda \subset \reals^{m}$ be an $n$-dimensional lattice with packing radius $\rho$.  Let $S$ be an $m$-dimensional hypersphere of radius $r$ centered at $t \in \reals^{m}$.  The number of lattice points from $\Lambda$ in the sphere $S$ is less than $\ceil{r/\rho}^{n}$ where $\ceil{\cdot}$ denotes the smallest integer strictly larger than its argument. 
 \end{proposition}
 \begin{proof}
The packing radius $\rho$ is the point on the boundary of the Voronoi cell $\vor(\Lambda)$ that is closest to the origin. Therefore, the sphere $S$ is a subset of $\vor(\Lambda)$ scaled by $r/\rho$ and translated by $t$.  That is, $S \subset r/\rho \vor(\Lambda) + t$.  The proof follows because the number of lattice points in $r/\rho \vor(\Lambda) + t$ is less than $\ceil{r/\rho}^n$ (this follows from, for example, \cite[Theorem~2]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}, see also~\cite[Lemma~3.7]{Micciancio09adeterministic}).
\end{proof}

@@

Proposition~\ref{obs:2} above asserts that after some finite number $K$ of iterations the procedure arrives at $x_K$, a closest lattice point to $y$.  Using Proposition~\ref{obs:1} we can detect that $x_K$ is a closest lattice point by asserting that $\|y - x_K\| \leq \| y - x_{K+1} \|$.
This simple iterative approach to compute a closest lattice point is related to what is called the \emph{iterative slicer}~\cite{Shalvi_iterativeslicer_2009}.  Micciancio~\cite{Micciancio09adeterministic} describes a related, but more sophisticated, iterative algorithm that can compute a closest lattice point in a number of operations that grows exponentially as $O(2^{2 n})$.  This single exponential growth in complexity is the best known.  %Most popular algorithms for computed closest lattice points require a number of operations of order $O(n^n)$\cite{Agrell2002,Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981}.

@@

@@

The root lattices $A_n$ and their dual lattices $A_n^*$ are classical examples of lattices of Voronoi's first kind~\cite{SPLAG}.  The construction of an `arbitrary' lattice of Voronoi's first kind appears nontrivial.  One approach would be to construct the matrix $Q = B^\prime B$ containing the Selling parameters $q_{ij}$.  The matrix $Q$ requires to have nonegative off diagonal entries and to be positive semidefinite with null space containing the vector $\onebf$ of all ones.  The explicit construction of such matrices is an interesting question for future research.  A related question is to determine, in some sense, how `large' the set of matrices satisfying these criteria is.  This would shed some light on how `large' the set of lattices of Voronoi's first kind is.

>>>>

\begin{proposition}\label{eq:latticepointsinsphere}
Let $\Lambda \subset \reals^{m}$ be an $n$-dimensional lattice with packing radius $\rho$.  Let $S$ be an $m$-dimensional hypersphere of radius $r$ centered at $t \in \reals^{m}$.  The number of lattice points from $\Lambda$ in the sphere $S$ is less than $\ceil{r/\rho}^{n}$ where $\ceil{\cdot}$ denotes the smallest integer strictly larger than its argument. 
 \end{proposition}
 \begin{proof}
The packing radius $\rho$ is the Euclidean length of a point on the boundary of the Voronoi cell $\vor(\Lambda)$ that is closest to the origin. Therefore, the sphere $S$ is a subset of $\vor(\Lambda)$ scaled by $r/\rho$ and translated by $t$.  That is, $S \subset r/\rho \vor(\Lambda) + t$.  The proof follows because the number of lattice points in $r/\rho \vor(\Lambda) + t$ is less than $\ceil{r/\rho}^n$ (this follows from, for example, \cite[Theorem~2]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}, see also~\cite[Lemma~3.7]{Micciancio09adeterministic}).
\end{proof}

@@

Proposition~\ref{obs:2} above asserts that after some finite number $K$ of iterations the procedure arrives at $x_K$, a closest lattice point to $y$.  Using Proposition~\ref{obs:1} we can detect that $x_K$ is a closest lattice point by checking whether $\|y - x_K\| \leq \| y - x_{K+1} \|$.
This simple iterative approach to compute a closest lattice point is related to what is called the \emph{iterative slicer}~\cite{Shalvi_iterativeslicer_2009}.  Micciancio~\cite{Micciancio09adeterministic} describes a related, but more sophisticated, iterative algorithm that can compute a closest lattice point in a number of operations that grows exponentially as $O(2^{2 n})$.  This single exponential growth in complexity is the best known.  %Most popular algorithms for computed closest lattice points require a number of operations of order $O(n^n)$\cite{Agrell2002,Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981}.

@@

% A number of important lattices are of Voronoi's first kind.  For example, the \emph{root lattices} $\ints^n, A_n, E_6, E_7$ and $E_8$ are of Voronoi's first kind~\cite{SPLAG}.  The root lattice $E_8$ has obtuse superbasis
% \[

% \]

% The root lattice $D_n$ has obtuse superbasis given by the vectors
% \begin{align*}
% b_1 &= -e_1 - e_2 \\
% b_i &= e_{i} - e_{i+1}, \qquad i = 2,\dots,n \\
% b_{n+1} &= e_{2} + e_{n}.
% \end{align*}

Classical examples of lattices of Voronoi's first kind are the $n$ dimensional root lattice $A_n$ and it's dual lattice $A_n^*$.  These lattices have convenient bases represented in $n+1$ dimensional space.  An obtuse superbasis for $A_n$ are $n+1$ vectors of length $n+1$ of the form
\begin{align*}
b_i &= e_i - e_{i+1}, \qquad i = 1, \dots, n \\
b_{n+1} &= e_{n+1} - e_{1},
\end{align*}
where $e_i$ is a vector of length $n+1$ with elements equal to zero except for the $i$th element that is equal to one.  The dual lattice $A_{n}^*$ is also of Voronoi's first kind.  The $i$th vector $b_i$ of the obtuse superbasis is of length $n+1$ with $j$th element equal to
\[
b_{ij} = \begin{cases}
\frac{n}{n+1} & j = i \\
-\frac{1}{n+1} & j \neq i.
\end{cases}
\]
For $A_n$ and $A_n^*$ there exists efficient algorithms that can compute a closest lattice point in $O(n)$ operations~\cite{McKilliam2009CoxeterLattices,Conway1982FastQuantDec}, and so, we do not recommend using the algorithm we describe here for these specific lattices.  %However, there exist a large number of lattice of first
%A lattice of Voronoi's first kind appears nontrivial.  One approach would be to construct the matrix $Q = B^\prime B$ containing the Selling parameters $q_{ij}$.  The matrix $Q$ requires to have nonpositive off diagonal entries and to be positive semidefinite with null space containing the vector $\onebf$ of all ones.  The explicit construction of such matrices is an interesting question for future research.  A related question is to determine, in some sense, how `large' the set of matrices satisfying these criteria is.  This would shed some light on how `large' the set of lattices of Voronoi's first kind is.

\section{A series of relevant vectors from a lattice of Voronoi's first kind}

@@
